---
title: "Project 1- Reproducible Research Week 2"
author: "Claudia Delgado"
date: "February 23, 2018"
output:
  html_document: default
---

##Loading and preprocessing the data
The data is organized in the csv file named is activity. Always verify that the file is in the working directory.
The following is the portion of the code for reading and processing the data. 
```{r Reading & Processing the data}
    getwd()
    stepdata <- read.csv("activity.csv", header=TRUE)
    colnames(stepdata)= c("steps", "date", "interval")
    summary(stepdata)
    stepdata$day <- weekdays(as.Date(stepdata$date))
    stepdata$DateTime<- as.POSIXct(stepdata$date, format="%Y-%m-%d")
    activity_no<- stepdata[!is.na(stepdata$steps),]
```

## Calculate the mean steps and plot a histogram

This portion of the code ignores missing values. 
Calculate the total number of steps taken per day.
Calculate and report the mean and median of the total number of steps taken per day. 
```{r Mean and Median}
stepsbydate <- aggregate(stepdata$steps~stepdata$date, FUN=sum,)
    colnames(stepsbydate)<- c("Date", "Steps")
    # Histogram of the total number of steps taken each day
    Figure1 <- hist(stepsbydate$Steps, ylab="Steps Frequency ", xlab="Number of Steps", main = "Total Steps Per Date", border="blue", 
         col="grey", breaks=5)
    summary(stepsbydate)
```
The mean and median were obtained by the functio summary or by the calcukation of each one separately as inicated in the R code. Th average number of steps taken each day was 10766 steps while the median number of steps taken each day was 10765 steps.


## What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r Time Series Plot }
    library(ggplot2)
    library(plyr)
    activity_no<- stepdata[!is.na(stepdata$steps),]
    intervals <- ddply(activity_no, .(interval), summarize, Avg = mean(steps))
    
    ##Create line plot of average number of steps per interval
    Figure2 <- ggplot(intervals, aes(x=interval, y=Avg), xlab = "Interval", ylab="Average Number of Steps")
    Figure2 + geom_line() + xlab("Interval")+ylab("Average Number of Steps")+ ggtitle("Average Number of Steps per Interval")
    summary(intervals)
    Max_Steps <- max(intervals$Avg)
    intervals[intervals$Avg==Max_Steps,1]
```

The 5-minute interval, on average across all the days in the dataset that contains the maximum number of steps is at the 835 interval. 

## Imputing missing values
``` {r Replacing missing values with mean}
sum(is.na(stepdata$steps))
    replacedata <- which(is.na(stepdata$steps))
    l <-length(replacedata)
    meansteps <-with(stepdata,tapply(steps,date,mean,na.rm=TRUE))
    nomissing <-mean(meansteps,na.rm = TRUE)
    for (i in 1:l) {
        stepdata[replacedata[i],1]<- nomissing
    }
    summary(replacedata)
    sum(is.na(replacedata))
    sum(is.na(stepdata))
    head(replacedata)
    steps<-with(stepdata,tapply(steps,date,sum,na.rm=TRUE))
    hist(steps, col="pink", breaks=5, xlab="daily steps", ylab="frequency",
         main="Total Number of Steps")
    summary(steps)
```
The total number of missing data (NA) was 2304.

The mean and median of total number of steps taken per day after imputing the data was 10770 steps.
Yes these values differ from the estimates from the first part of the assignment.
The impact of imputing missing data on the estimates of the total daily number of steps is minimum.

## Are there differences in activity patterns between weekdays and weekends?
For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

``` {r weekdays vs weekends}
library(plyr)
    library (ggplot2)
    library (lattice)
    stepdata$DayCategory <- ifelse(stepdata$day %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
    weekday <- ddply(stepdata, .(interval, DayCategory), summarize, Avg = mean(steps))
    xyplot(Avg~interval|DayCategory, data=weekday, type="l",  layout = c(2,1),
           main="Intervals Step Average by Day Type", 
           ylab="Steps Average", xlab="Interval")
    
```

Yes there are differences in the intervals between weekdays and weekends
